Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-11-08 manuscript/manuscript.Rmd
Housing data

To be continued.

@@

```
We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic. Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}. Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80. This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}, which derive from the type I error rates of the Yeo-Johnson transformation.

@@

```{=latex}
\begin{table}
\begin{center}
\caption{Test statistic for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
\label{tab:empirical-central-normality}
\begin{tabular}{l c c c c c c c}

\toprule
transformation & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
rob. shift-sensitive Box-Cox     & 0.046 & 0.072 & 0.090 & 0.106 & 0.126 & 0.140 & 0.189 \\
rob. shift-sensitive Yeo-Johnson & 0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

@@

In this section, we assess shift sensitive power transformations using real datasets.

@@

```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
cap <- paste0(
  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
  "This data contains three strong outliers for highly efficient vehicles. ",
  "The left panel shows the quantile-quantile plot for the original data. ",
  "The middle panel shows the quantile-quantile plots for data transformed using ",
  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2021-kq]. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
  "The right panels shows the quantile-quantile plots for data transformed using ",
  "the non-robust and robust shift-sensitive transformations. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

.plot_top_gear(
  plot_theme = plot_theme
)

@@

>>>>

@@

```
We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic. Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}. Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80. In the remaining we will use the empirical central normality test statistic values defined using robust shift-sensitive Yeo-Johnson transformations, as it is slightly more strict. This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}

@@

```{=latex}
\begin{table}
\begin{center}
\caption{Test statistic for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
\label{tab:empirical-central-normality}
\begin{tabular}{c c c c c c c}

\toprule
0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

@@

In the previous section, we simulated data to assess shift-sensitive power transformations and their robustness against outliers. In this section, we assess these aspects using real datasets.

@@

```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
cap <- paste0(
  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
  "This data contains three strong outliers for highly efficient vehicles. ",
  "The left panel shows the quantile-quantile plot for the original data. ",
  "The middle panel shows the quantile-quantile plots for data transformed using ",
  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \cite{Raymaekers2021-kq}. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
  "The right panels shows the quantile-quantile plots for data transformed using ",
  "the non-robust and robust shift-sensitive transformations. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

.plot_top_gear(
  plot_theme = plot_theme
)
```

@@

```{r housing-data, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
cap <- paste0(
  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
  "This data contains three strong outliers for highly efficient vehicles. ",
  "The left panel shows the quantile-quantile plot for the original data. ",
  "The middle panel shows the quantile-quantile plots for data transformed using ",
  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2021-kq]. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
  "The right panels shows the quantile-quantile plots for data transformed using ",
  "the non-robust and robust shift-sensitive transformations. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

.plot_sacramento_house_prices(
  plot_theme = plot_theme
)
```
